---
title: "Statistical Inference Course Project Part 1"
author: "Bamini Balaji"
date: "January 22, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Simulation


### Overview

The objective is to simulate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.  


### Simulation Code

Let's set lambda = 0.2 for all of the simulations.

First, we will simulate 1000 sets of 40 exponentials. Then, we will compute their means and store it in a vector of length 1000 named "means".


```{r}
lambda <- 0.2
n = 40 # This indicates number of distributions to be averaged
s = 1000 # This is number of simulations
sim_data <- matrix(rexp(s*n, lambda), nrow = n, ncol = s)
means <- apply(sim_data, 2, mean)
sdev <- apply(sim_data, 2, sd)
avgmeans <- mean(means)
avgsd <- mean(sdev)
```


### Analysis of Mean

The average of ovserved means is `r avgmeans`. 

The theoretical mean of this distribution is 1/lambda or `r 1/lambda`.

The following figure shows the observed distribution of means through a histogram. The expected value of the theoretical mean is shown by the red line.  

This looks like a normal distribution!

```{r, echo=FALSE}
means <- data.frame(means)
library(ggplot2)
g <- ggplot(means, aes(means))
g + geom_histogram(binwidth = 0.3, color = "black", aes(y = ..density..)) + geom_vline(xintercept = 1/.2, color = "red", lwd = 1.2) 
```

Let's perform a hypothesis test with alpha of 0.05:

* H_0 : mu = 5.0
* H_a : mu != 5.0

```{r}
t.test(means, mu = 5, conf.level = 0.95)
```

Null hypothesis is rejected if p-value is less than 0.05 (alpha).


### Analysis of Variance

The measured average standard deviation equals `r avgsd`.

The theoretical standard deviation of an exponential distribution is 1/lambda or `r 1/lambda`.

The standard deviation of the observed distribution is theoretically sigma/sqrt(n), or `r 1/lambda/sqrt(40)` where n = 40 and sigma is 1/lambda. 

Let's overlay a normal distribution to this figure. 
Z ~ N(5, 5/sqrt(40))

This distribution is illustrated using the blue curve.


```{r, echo=FALSE}
g + geom_histogram(binwidth = 0.3, color = "black", aes(y = ..density..)) + geom_vline(xintercept = 1/.2, color = "red", lwd = 1.2) + stat_function(fun = dnorm, args = list(mean = 5, sd = 1/lambda/sqrt(40)), color = "blue", lwd = 1)
```

Let's perform a hypothesis test with alpha of 0.05:

* H_0 : sd = 5.0
* H_a : sd != 5.0

```{r}
t.test(sdev, mu = 5, conf.level = 0.95)
```

Null hypothesis is rejected if p-value is less than 0.05 (alpha). 

